Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-22 trunk/doc/thesis/AUI/report_AUI.tex
MODIFICATIONS DONE! YA RABBI

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
\usepackage[titletoc]{appendix}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\usepackage{etoolbox}
\makeatletter
\patchcmd{\thebibliography}{%
  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
  \chapter{References}}{}{}
\makeatother

\begin{document}
\doublespacing
\input{acronym}
\frontmatter

%COVER PAGE
\input{cover_page}
\chapter{} % dedications chapter
\begin{dedication}
To my dear parents for their patience, love and support \\
To my brothers Hossein, Omar and Saad \\
To my cousin Sara and her family \\
To my family \\
To my friends \\

\end{dedication}
\chapter{Acknowledgments}
\paragraph{}
I am deeply thankful to my supervisor Dr. Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project. Your advices whether they are computer science related, or even every day life related, are wise words that I will never forget.
\paragraph{}
I am also thankful for having Dr. Driss Benhaddou as a supervisor. His advice, insight, guidance and very enlightening ideas. I wouldn't have been so interested in smart grids without the discussions we had on the topic.
\paragraph{}
I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
\paragraph{}
I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
\paragraph{}
My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
\paragraph{}
I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.

\chapter{Abstract}
\paragraph{}
Information Technology is growing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. It is smoothly evolving from an Internet of People to and Internet of \textit{Things}. By 2020, the Internet is expected to have 50 billion \textit{Things} connected to it. 
\paragraph{}
However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
Another migration is taking place in the power grid that is evolving towards the \gls{sg}. In fact, Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}. The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
\paragraph{}
This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for Smart Grid applications. The test-bed built provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}. Therefore, the system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.

\chapter*{Résumé}
\paragraph{}
Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies. L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}. L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un réseau pour les \textit{Choses}. En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
\paragraph{}
Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
\paragraph{}
Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}. Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et l'exploitation décentralisée des énergies renouvelables. Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
\paragraph{} 
Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid. Ce prototype fournit des services tels qu'une passerelle IPv6, le contrôle à distance des capteurs sans fil, et un accès mobile au système.
\paragraph{}
Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}. Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype. Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
\setcode{utf8}

@@

\mainmatter
\doublespacing
\chapter{Introduction}

@@

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. Such an integration through the two-way communication will to make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not doable. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. The use of different technologies comes at a cost which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed with the two-way communication. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. To solve such a problem, we add an extra component that will track all  packets and change them to fit the standards of the targeted component.

\section{Scope of Study}
\paragraph{}
A variety of tools and technologies were used for this study.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that will be integrated into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

\section{Outline of the Thesis}
\paragraph{}
The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context. This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ. Finally, its presents the commonalities ad differences between this thesis and the work previously done.
\paragraph{Chapter 3}
This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
\paragraph{Chapter 4}
It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter mostly discusses the system deployment. It starts by presenting the hardware used in this project, then moves to the technology enablers. Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
\paragraph{Chapter 6}
It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
\paragraph{Chapter 7, 8}
These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}. Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer. Households can now produce and sell the surplus of electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it. Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} is divided into three layers:

@@

\paragraph{}
From figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer. The first one is the \gls{ipv4}/\gls{ipv6} gatewaying process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them in order to meet the recipient's communication technology and standards. The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel; This component is in charge of receiving raw \gls{ipv6} packets and encapsulate them in  USB frames to reach the destination. The third component is deployed in the sink which is the entry and exit point of the \gls{wsn}. This special mote receives \gls{ipv6} packets encapsulated in USB frames that are forwarded to the destination mote using Zigbee.

>>>>

\documentclass[12pt,a4paper,final]{report}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
\usepackage[titletoc]{appendix}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\usepackage{etoolbox}
\makeatletter
\patchcmd{\thebibliography}{%
  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
  \chapter{References}}{}{}
\makeatother
\usepackage{chngcntr}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

\begin{document}
\doublespacing
\input{acronym}

%COVER PAGE
\input{cover_page}
\chapter*{} % dedications chapter
\begin{dedication}
To my dear parents for their patience, love and support \\
To my brothers Hossein, Omar and Saad \\
To my cousin Sara and her family \\
To my family \\
To my friends \\
\end{dedication}

% display page numbers in the headings. Start with roman numerals %
\pagestyle{headings}
\setcounter{page}{1}
\pagenumbering{roman}

\chapter*{Acknowledgments}
\paragraph{}
I am deeply thankful to my supervisor Dr. Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project. Your advices whether they are computer science related, or even every day life related, are wise words that I will never forget.
\paragraph{}
I am also thankful for having Dr. Driss Benhaddou as a supervisor. His advice, insight, guidance and very enlightening ideas. I wouldn't have been so interested in smart grids without the discussions we had on the topic.
\paragraph{}
I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
\paragraph{}
I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
\paragraph{}
My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
\paragraph{}
I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.

\chapter*{Abstract}
\paragraph{}
Information Technology is growing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}. By 2020, it is expected to have 50 billion \textit{Things} connected to it. 
\paragraph{}
However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
Another migration is taking place in the power grid that is evolving towards the \gls{sg}. In fact, Information Technology (IT) is playing a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}. The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
\paragraph{}
This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for Smart Grid applications. The test-bed provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
The project contributes to SG by settling the IT infrastructure  (i.e. Network and protocol) that is needed to disseminate information the SG. This infrastructure allows a remote monitoring and control of appliances connected to such infrastructure and therefore the two-way communication needed by SG.
\paragraph{}
This study leads to the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}. The system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.

\chapter*{Résumé}
\paragraph{}
Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies. L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}. L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un réseau pour les \textit{Choses}. En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
\paragraph{}
Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
\paragraph{}
Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}. Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et l'exploitation décentralisée des énergies renouvelables. Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
\paragraph{} 
Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid. Ce prototype fournit des services tels qu'une passerelle IPv6, le contrôle à distance des capteurs sans fil, et un accès mobile au système.
\paragraph{}
Le projet contribue à SG en construisant l'infrastructure IT (ex. réseau et protocole) qui est nécessaire pour diffuser l'information au SG. Cette infrastructure permet une surveillance et le contrôle à distance des appareils connectés à cette infrastructure et donc la communication bidirectionnelle nécessaires pour SG.
\paragraph{}
Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}. Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype. Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
\setcode{utf8}

@@

\doublespacing
\chapter{Preamble - Thesis Research Methodology}
\setcounter{page}{1}
\pagenumbering{arabic}
\paragraph{}
 In the course of this study, a set of rules and standards dictated the research methodology. With the adoption of this methodology and mainly the scientific method, the study resulted in a test-bed that is presented in this report along with the experiments conducted and measurements collected. This chapter will briefly explain the phases through which the study passed through and how it influenced the different directions taken throughout the course of this project.
\section{Introduction}
\paragraph{}
Everything started with an idea: to be able to monitor and control appliances along with integrating it in the Smart Grid. As an idea, many things are very abstract at this point. the what question is not clear: What are we trying to build? What are we trying to get at the end? What should it do? Afterwards comes the how questions: How are building such a system? How can we monitor and control appliances? How should it be part of smart Grids? Such questions lead to the second step in this research methodology which getting to look at the state of the art, how such things are done nowadays and what already exists in the market.

\section{State of the Art}
In this step, the goal is to get acquainted with the main concepts that are used in this study such as Smart Grid and the Internet of Things. To do so, a review of the work done in these two fields has been performed. Such review includes reading the various papers from different journals, learning more about the projects that are in development, their approaches, the existing architectures, the standards adopted and the tools used.
The academic perspective gives a deep understanding of the topic, the options, issues, challenges and outcomes that exist but there is a need to see what industry is actually building. As a matter of fact, a research has been conducted to have an idea about the current commercial applications that exist nowadays, how are they built and what do they perform. This step gives us a clear idea about what has been achieved both in academia and industry.
Now that we have a clear understanding of the main concepts and tools that exist, it is time to start bringing our contribution to the field but before that, the objectives should be clearer and more accurate than the initial proposed idea at the beginning of the study
\section{Development and Analysis}
The problem statement is restated as well the purpose and scope of the study. It is based on the initial objectives but takes into consideration numerous ideas, issues and challenges that arise from looking at the state of the art.
An initial architecture is sketched, that shows the main blocks of the work. After building such architecture comes different directions and one should choose and defend the most appropriate for this study. In this study, numerous alternatives were raised (e.g. whether the wireless sensor network should be mesh-based or tree-based) and we had to pick the most appropriate one with respect to our objectives and constraints namely time and cost.
Afterwards, in each component that have to be developed, numerous directions and alternatives appeared, and again the most appropriate that fulfils our needs was chosen as it will be discussed in this thesis. The rest of the development followed the same trend which is an iterative development model but at the same time modular. Each component, was studied separately where different research decisions were taken and as it was divided into several subcomponents, we studied each of them separately. The interconnection between different components made this work iterative because each decision in a component affects the rest of the components and therefore this decision had to be studied to be aware of its effect on the rest of the system. For example, when we decided to have a TCP client in the mote, this automatically affected the gateway component, because there had to be a TCP server there too. If UDP was chosen in the mote, this decision would have affected the gateway component too.
This approach was followed until the whole system was developed and analysed. Once the system was ready, the next step is to see if is solves the problem stated in the first step and if it fulfils the outcomes of this study
\section{Measurements of the Results}
In this step, the goal is to check if the system fulfils its requirements. To do so, an initial testing phase was performed where the system was put to test. The functional requirements were tested extensively in order to identify the main problems the system could suffer from.
Afterwards, the system's various non functional requirements were put to test. The first one is to check is to check if the system could host overlying services. To do so, it should meet some performance qualities and criteria. As a matter of fact, these performance criteria were identified based on measurement and past studies. Once these criteria have been identified, the system's performance had to tested. To do so, experiments were conducted to be able to collect data that can show whether the system has met these performance criteria or not.
After these experiments has been conducted, the next step is to report the results of the this study which includes all the steps through which the study through.

\section{Report}
In this step, a report has been written showing all the steps through which the study has passed through. This includes the introduction, the review of the state of the art, the development and analysis and the measurements of the result. The report is a concise document showing all the information, alternatives, decisions and outcomes dealt with throughout the course of this study.

\chapter{Introduction}

% set the page numbers to be arabic, starting at page 1 %

@@

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. Such an integration through the two-way communication will make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely. Such pillar is absent for the time being, and therefore the integration of the smart home within the smart grid is not doable. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. The use of different technologies comes at a cost which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed with the two-way communication. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \gls{iot}. To solve such a problem, we add an extra component that will track all  packets and change them to fit the standards of the targeted component.

\section{Scope of Study}
\paragraph{}
A variety of tools and technologies were used for this study.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from the deployment perspective in Chapter 5. The reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that will be integrated into the \gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

\section{Outline of the Thesis}
The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 3}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ. Finally, its presents the commonalities ad differences between this thesis and the work previously done.
\paragraph{Chapter 4}
This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
\paragraph{Chapter 5}
It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
\paragraph{Chapter 6}
This chapter mostly discusses the system deployment. It starts by presenting the hardware used in this project, then moves to the technology enablers. Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
\paragraph{Chapter 7}
It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
\paragraph{Chapter 8, 9}
These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.
\section*{}
Now that the problem statement has been defined in addition to the purpose of the study and the scope, a review of the current work and the state of the art will be discussed in chapter 3 and 4.

@@

The current electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to a new technology: the \gls{sg}. Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer. Households can now produce and sell the surplus of electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it. Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favour  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} is divided into three layers:

@@

\paragraph{}
From Figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer. The first one is the \gls{ipv4}/\gls{ipv6} gateway process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them in order to meet the recipient's communication technology and standards. The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel; this component is in charge of receiving raw \gls{ipv6} packets and encapsulate them in  USB frames to reach the destination. The third component is deployed in the sink which is the entry and exit point of the \gls{wsn}. This special mote receives \gls{ipv6} packets encapsulated in USB frames that are forwarded to the destination mote using Zigbee.
